---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
  md_document:
    variant: markdown_github
---

```{r, echo=FALSE}
library(ggplot2)
library(knitr)
```

## Rotterdam2 pre-QC
Initial quality assessment of dataset clustered by HUGE-F in Rotterdam. 
```{r,echo=FALSE}
# Load sample sheet

samples <- read.table('/home/oyvind/hunt-cloud/mnt/archive/ROTTERDAM2/data/GSA2017_246_025_ReRun/GSA2017_246_025_STS.csv', skip = 9, header = T, stringsAsFactors = F, sep = ',')

rawpath = '/home/oyvind/hunt-cloud/mnt/archive/ROTTERDAM2/qc/qc-rotterdam2-out/pre-qc-rotterdam2-output/'

het <- read.table(gzfile(file.path(rawpath, 'raw-het.het.gz')), header = T, stringsAsFactors = F) 
imiss <- read.table(gzfile(file.path(rawpath, 'raw-missing.imiss.gz')), header = T, stringsAsFactors = F) 

#lmiss <- read.table(gzfile(file.path(rawpath, 'raw-missing.lmiss.gz')), header = T, stringsAsFactors = F) 
#frq <- read.table(gzfile(file.path(rawpath, 'raw-frq.frq.gz')), header = T, stringsAsFactors = F) 
#hwe <- read.table(gzfile(file.path(rawpath, 'raw-hwe.hwe.gz')), header = T, stringsAsFactors = F)   

samples.het <- merge(samples, het, by.x='Sample_ID', by.y='IID')
samples.missing <- merge(samples, imiss, by.x='Sample_ID', by.y='IID')
```

```{r, echo=F}
# Function for plotting mean Y binned by X
mean.batch.plot <- function(data, x,y, labx, laby, title) {
  p <- ggplot(data, aes(factor(get(x)), get(y))) + 
  geom_boxplot() + 
  stat_summary(fun.data=mean_cl_boot, geom="errorbar", colour='red') +
  stat_summary(fun.y=mean, geom="point", colour='red') + theme(axis.text.x = element_text(size=rel(0.65), angle = 90, hjust = 1)) +
  ggtitle(title) +
  xlab(labx) +
  ylab(laby)
  
  return(p)
}
```

### Call rate  

Overall call rate per sample.  

```{r, echo=F}
samples.missing$call_rate <- 1- samples.missing$F_MISS
samples.missing$sample <- factor(samples.missing$Sample_ID, levels = samples.missing$Sample_ID[order(samples.missing$call_rate)])
ggplot(samples.missing) + geom_point(aes(sample, call_rate))  
ggplot(samples.missing) + 
  geom_point(aes(sample, call_rate)) + 
  scale_y_continuous(limits = c(0.85, 1), breaks=seq(0.85,1,0.01)) + 
  geom_hline(yintercept = 0.98, colour='red')

samples.missing$valcut <- cut(samples.missing$call_rate, breaks=c(0, 0.8, 0.9, 0.95,0.96,0.97, 0.98, 0.99, 0.995, 1))
t <- as.data.frame(table(samples.missing$valcut))
names(t) <- c('Call rate','N')
kable(t)
```

### Heterozygosity
Mean heterozygosity per:

- Sample plate
- Sample well
- Sentrix Barcode/ID
- Sentrix position

```{r, echo=F}
mean.batch.plot(samples.het, 'Sample_Plate', 'F', 'Sample plate', 'Heterozygosity F-stat', "Sample plate vs. mean heterozygosity rate")
mean.batch.plot(samples.het, 'Sample_Well', 'F', 'Sample well', 'Heterozygosity F-stat', "Sample well vs. mean heterozygosity rate")
mean.batch.plot(samples.het, 'SentrixBarcode_A', 'F', 'Sentrix barcode', 'Heterozygosity F-stat', "Sentrix ID vs. mean heterozygosity rate")
mean.batch.plot(samples.het, 'SentrixPosition_A', 'F', 'Sentrix position', 'Heterozygosity F-stat', "Sentrix position vs. mean heterozygosity rate")
```

### Call rate
Mean call rate per: 

- Sample plate
- Sample well
- Sentrix Barcode/ID
- Sentrix position

```{r, echo=F}
mean.batch.plot(samples.missing, 'Sample_Plate', 'F_MISS', 'Sample plate', 'Sample missing freq.', "Sample plate vs. mean missing frequency")
mean.batch.plot(samples.missing, 'Sample_Well', 'F_MISS', 'Sample well', 'Sample missing freq.', "Sample well vs. mean missing frequency")
mean.batch.plot(samples.missing, 'SentrixBarcode_A', 'F_MISS', 'Sentrix barcode', 'Sample missing freq.', "Sentrix ID vs. mean missing frequency")
mean.batch.plot(samples.missing, 'SentrixPosition_A', 'F_MISS', 'Sentrix position', 'Sample missing freq.', "Sentrix position vs. mean missing frequency")
```



### Call rate vs. heterozygosity
Plots showing relationship between call rate and heterozygosity F-stat. Higher missingness yields more heterozygous calls. 

```{r, echo=F}
samples.missing.het <- merge(samples.missing, het, by.x='Sample_ID', by.y='IID')

ggplot(samples.missing.het) + 
  geom_point(aes(F, F_MISS)) + 
  xlab("Heterozygosity F-stat") +
  ylab("Frequency of missing calls")

ggplot(samples.missing.het) + 
  geom_point(aes(F, F_MISS)) + 
  xlab("Heterozygosity F-stat") +
  ylab("Frequency of missing calls") +
  coord_cartesian(ylim = c(0,0.03), xlim=c(-0.25,0.25))

```

